---
title: "Salmon lab"
author: "An Bui"
date: "21 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##### 1-3. Set up
```{r set up packages and csv, include = FALSE}
library(tidyverse)
library(kableExtra)

salmon <- read_csv("salmon_imports.csv")
```

##### 4. Exploratory scatterplot
```{r exploratory qqplot, include = FALSE}
salmon_qq <- ggplot(salmon, aes(sample = salmon_fresh)) +
  geom_qq()

salmon_qq
# seems linear
```

```{r exploratory scatterplot, echo = FALSE}
salmon_scatter <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point()

salmon_scatter
# generally linear, with some outliers? linear regression still makes sense though
```

##### 5. Linear regression in R
```{r linear regression}
# do a linear regression of pounds of salmon per year from data frame salmon
salmon_model <- lm(salmon_fresh ~ year, data = salmon)

salmon_model
```
Model equation: imports(tons) = 7550x + -14982940  
Slope: 7550 (what does this mean??)  
Intercept: At year 0, -14972940 pounds of salmon will be harvested a year. This is obviously unrealistic, so this model is not the best fit for extrapolating past values.  

##### 6. Model diagnostics
```{r model diagnostics, echo = FALSE}
par(mfrow = c(2,2))

plot(salmon_model)
```

- Residuals vs Fitted: looks homoscedastic  
- QQ plot: residuals are normally distributed  
- Scale-Location: how to explain?  
- Residuals vs Leverage: points 24 and 28 are influencing the model  

##### 7. Explore model fit and significance
```{r model summary}
summary(salmon_model)
```

- _Does year significantly predict salmon imports?_ Year predicts 94% of variance in salmon imports (_p_ < 0.001).  
- _What does the R^2^ value actually mean in words?_ 94% of variance in salmon imports is predicted by year.  
- _What proportion of the variation in salmon imports is NOT explained by year?_ 6% of the variation in salmon imports is not explained by year.  

##### 8. Prepare a final graph. 
```{r final graph, echo = FALSE}
final_x = "Year"
final_y = "US Annual Fresh Salmon Imports (1 000 pounds)"

final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point() +
  geom_smooth(method = lm, 
              se = TRUE, 
              size = 0.5, 
              color = "gray20") +
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  labs(x = final_x, y = final_y)

final_graph

```

##### 9. Make predictions using the model
What will salmon imports look like in 2022, 2024, 2026, and 2028?
```{r predictions, include = FALSE}
# make data frame of with sequence of years that you want to make predictions for
# write in column name that is the same as the variable name 'year' that the model uses
new_years <- data.frame(year = c(2022, 2024, 2026, 2028))

# use predict() to create table of predicted values
future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")

future_predict

# bind prediction outcomes with new_years data
predictions <- data.frame(new_years, future_predict)

predictions
```

```{r table, echo = FALSE}
# put a table on it
table_caption = "**Table 1. Salmon import predictions.** Predictions of salmon imports (1 000 pounds) using linear mixed model for 2022, 2024, 2026, and 2028. Limits are within ± 95% CI."
table_columns = c("Year", "Predicted value", "Lower limit", "Upper limit")

kable(predictions,
      col.names = table_columns,
      caption = table_caption) %>% 
  kable_styling(bootstrap_options = "striped")
```

##### 10. Find Pearson's _r_ using cor.test()
```{r find r, include = FALSE}
salmon_cor <- cor.test(salmon$year, salmon$salmon_fresh)

salmon_cor
```
There is a strong positive correlation between year and salmon imports (1 000s pounds, Pearson's _r_ = `r round(salmon_cor$estimate, 2)`). The model (_salmon imports_ = 7550 * (_year_) + -14982940) explains a significant amount of variance in salmon imports (F(`r salmon_model$df.residual`) = 471.9, _p_ < 0.001, _R^2^_ = 0.95).
















